\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.72in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\usepackage{microtype}
\usepackage{enumitem}

\setlist[itemize]{leftmargin=1.4em,itemsep=0.12em,topsep=0.2em}
\setlength{\parskip}{0.32em}
\setlength{\parindent}{0pt}
\pagestyle{empty}

\begin{document}

\begin{flushright}
Mehmet Ergezer\\
Wentworth Institute of Technology\\
550 Huntington Avenue\\
Boston, MA 02115, USA\\
\today
\end{flushright}

Editors\\
\textit{Mathematical Programming Computation}

Dear Editors,

Please consider the manuscript
\textit{``Salem--Spencer sets as a cross-solver benchmark: decomposition,
certification, and split-policy effects at $r_3(212)$''}
for publication in \textit{Mathematical Programming Computation}.

The manuscript presents a reproducible computational study and benchmark
family built from the open frontier decision problem for OEIS A003002. It does
not claim to resolve $r_3(212)$. Instead, it develops and evaluates a
witness-informed cube decomposition, quantifies the effect of
problem-specific window-cardinality cuts, compares CP-SAT, LP/MIP, and CDCL
formulations, and releases a graded instance library with machine-checkable
proof artifacts.

The study contains five results aligned with MPC's emphasis on practical
computation, comparative testing, reusable software, and verification:

\begin{itemize}
\item Six monolithic 24-hour baselines fail. A matched five-policy ablation
  exposes a cover-size versus conquer-cost tradeoff: global AP-degree reduces
  the exact emitted cover from 12,582,912 to 96,847 cubes, but leaves
  individually harder subproblems under the historical CP-SAT cap. A
  standalone C implementation independently reproduces the complete cover.
\item A two-stage CDCL survey closes 6,045 of 6,071 broad residuals (99.57\%)
  with no feasible result, while 26 released instances remain as a sharply
  defined challenge tier.
\item On 20 byte-identical hard-pocket formulas run sequentially on matched
  AMD EPYC 7763 nodes, CaDiCaL 3.0.0 and kissat 4.0.4 both close every
  instance. Kissat wins all 20 pairs, with geometric-mean CaDiCaL-to-kissat
  time ratio 1.477 (bootstrap 95\% interval 1.332--1.655).
\item Six solve-time-stratified survivors from the alternative global-degree
  cover reproduce the survey CNF hashes, verify through DRAT-to-LRAT
  conversion, and are accepted by the formally verified \texttt{cake\_lpr}
  checker. Separate end-to-end regressions recover the known exact values
  $r_3(80)=22$, $r_3(90)=24$, and $r_3(100)=27$ from independently verified
  lower witnesses and \texttt{cake\_lpr}-checked upper certificates.
\item Optimization-form HiGHS experiments show that window inequalities
  reduce the T1b node count by approximately 81\% and tighten every tested LP
  and MIP upper bound to the decision threshold 44, but not below it.
\end{itemize}

Source code and benchmark generators are available at
\url{https://github.com/memoatwit/erdos_1194/tree/mpc-submission-v1.3/erdos_r3}.
Versioned artifacts are archived at
\url{https://doi.org/10.5281/zenodo.21413746}. The
journal-submission version of the archive includes formulas, solver outputs,
proof objects, hashes, an independent model audit, and machine-readable
provenance, including the six global-degree certificate bundles. The preprint
is available as arXiv:2606.04016. The manuscript is not under consideration at
another journal.

For Technical Editor review, \texttt{ARTIFACT\_REPRODUCIBILITY.md} maps each
claim to its data and reproduction command. The $N=80$ exact-value regression
is the fastest end-to-end route and runs in minutes on a workstation; the
archived certificates can also be checked with \texttt{drat-trim} and
\texttt{cake\_lpr} without re-solving the instances.

The paper retains and explicitly refutes a working hypothesis formed during
the campaign. We believe this is methodologically useful: it demonstrates why
unresolved instances, solver configurations, and proof objects should be
archived and retested under controlled pipelines.

Generative-AI use is disclosed in the manuscript. The author made all final
scientific and experimental decisions, independently checked the numerical
and mathematical claims, and takes full responsibility for the work.

Sincerely,

\vspace{0.8em}
Mehmet Ergezer\\
Wentworth Institute of Technology\\
\href{mailto:ergezerm@wit.edu}{ergezerm@wit.edu}\\
ORCID: \href{https://orcid.org/0000-0001-6627-3667}{0000-0001-6627-3667}

\end{document}
