% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\section{Cover letter --- Experimental
Mathematics}\label{cover-letter-experimental-mathematics}

Mehmet Ergezer\\
Wentworth Institute of Technology\\
550 Huntington Avenue\\
Boston, MA 02115, USA\\
ergezerm@wit.edu\\
ORCID: 0000-0001-6627-3667

June 10, 2026

The Editors\\
Experimental Mathematics

Dear Editors,

I am pleased to submit the manuscript \emph{``Witness-split +
window-cardinality\\
refinement for r\_3(N): architecture, empirical results, and a
structural hard\\
pocket''} for consideration in Experimental Mathematics. The preprint is
posted\\
at arXiv:2606.04016 (https://arxiv.org/abs/2606.04016). All campaign
artifacts,\\
including verified DRAT/LRAT proofs and the Lean encoding of the
residual\\
problem, are deposited at Zenodo (DOI 10.5281/zenodo.21413746).

The paper describes a reproducible computational framework for
upper-bound\\
search on r\_3(N), the maximum size of a
3-term-arithmetic-progression-free\\
subset of {[}1, N{]}. Applied to the OEIS A003002 frontier case N = 212,
K = 44,\\
the campaign found no feasible 44-element 3-AP-free subset of {[}1,
212{]} across\\
millions of CP-SAT subproblems, providing strong empirical support for
the\\
conjectural value r\_3(212) = 43 without producing a formal proof. The
technical\\
contribution is methodological:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  A witness-split + window-cardinality CP-SAT architecture that is\\
  parametric in (N, K) and reusable for r\_3(N') at adjacent N.
\item
  A characterization of the structural hard pocket of the residual
  that\\
  resists CP-SAT propagation, recursive refinement, redundant Tseitin\\
  clauses, wall-cap extension, and HiGHS LP-relaxation MIP. A
  subsequent\\
  CDCL/SAT attack closes 18 of the 20 LP-paradigm-resistant chunks,
  all\\
  with independently verified DRAT proofs (verified by drat-trim).
\item
  A genuinely paradigm-resistant residual T1c, of size 2, that
  survives\\
  CP-SAT, HiGHS, pure CDCL at a 12-hour wall, and CDCL with windowed\\
  cardinality at a 4-hour wall. T1c is released both as JSONL solver\\
  instances and as a Lean 4 formal-proof-search encoding compatible
  with\\
  AlphaProof Nexus-style EVOLVE-BLOCK conventions.
\end{enumerate}

I believe Experimental Mathematics is the natural venue for this work.
The\\
campaign is squarely within the journal's stated scope: an experimental
and\\
reproducibility-focused investigation of a frontier integer-sequence
value,\\
with a release-grade dataset of verified machine-checkable certificates,
an\\
honestly negative methodological result (no value of r\_3(212) is
proved), and\\
a precise open problem framed for follow-on attack. The paper would
benefit\\
from review by specialists in additive combinatorics, constraint
solving, and\\
formal verification.

The work is not under consideration at another journal. The preprint\\
arXiv:2606.04016 was posted to establish a citable timestamp; this is
standard\\
practice for the field and consistent with Experimental Mathematics'
position\\
on preprints.

\section{Disclosure of generative-AI
use}\label{disclosure-of-generative-ai-use}

In accordance with the Taylor \& Francis AI Policy:

Generative AI tools (Anthropic Claude Sonnet 4.5 / Claude Opus 4.6,
accessed\\
via the Claude Code agent interface during 2026; and OpenAI Codex,
GPT-5,\\
accessed via the Codex agent interface during 2026) were used during
the\\
preparation of this manuscript. The use was as follows:

\begin{itemize}
\item
  Drafting and revision assistance for the manuscript sections. The\\
  author provided the campaign data, the technical results, the
  empirical\\
  interpretation, and the headline framing decisions. The AI tool was
  used\\
  to draft initial section text from author-provided notes, propose\\
  structural improvements (subsection ordering, transition phrasing,\\
  benchmark-tier organization), and run consistency sweeps for numbers\\
  and cross-references across drafts. Every claim in the final
  manuscript\\
  was reviewed, verified, and either accepted, edited, or rejected by
  the\\
  author.
\item
  Drafting assistance for repository scaffolding (build scripts,\\
  bibliography seed, LaTeX driver, ancillary-file packaging),
  including\\
  the preparation of journal-submission support files. All code was\\
  reviewed and the build outputs were verified by the author before\\
  submission.
\item
  Computational orchestration assistance during the campaign (Unity\\
  SLURM scripts, results aggregation, proof-verification wrappers).\\
  The mathematical content of the campaign and all solver-paradigm,\\
  encoding, and bound design decisions were made by the author.
\end{itemize}

The AI tools did not generate mathematical claims, did not select or run
the\\
underlying computational experiments without author direction, and did
not\\
serve as authors. All mathematical conclusions, the campaign design,
the\\
formal-conjecture framing, and the manuscript's central scientific
argument\\
are the author's own.

The author confirms that the manuscript complies in full with the Taylor
\&\\
Francis AI Policy.

\section{Suggested reviewers}\label{suggested-reviewers}

The following specialists are well-positioned to review the manuscript:

\begin{itemize}
\item
  Thomas F. Bloom, University of Manchester\\
  (expert on density-increment methods for r\_3 and Erdős upper-bound\\
  problems; maintains erdosproblems.com).
\item
  Marijn J. H. Heule, Carnegie Mellon University\\
  (expert on DRAT/LRAT proof certificates and drat-trim; would be\\
  well-placed to assess the verification chain).
\item
  Lorenzo Cariboni\\
  (maintains the OEIS A003002 b-file at the current frontier; the
  natural\\
  expert on the prior numerical record).
\item
  Olof Sisask, Stockholm University\\
  (additive combinatorics; co-author of the Bloom-Sisask
  density-increment\\
  framework cited in the manuscript).
\end{itemize}

\section{Data and reproducibility}\label{data-and-reproducibility}

All data, code, verified proofs, and the Lean formalization required
to\\
reproduce the campaign and verify every certificate-track claim are
available\\
under open license at:

\begin{itemize}
\tightlist
\item
  Preprint: arXiv:2606.04016
\item
  Dataset: 10.5281/zenodo.21413746 (CC-BY)
\end{itemize}

I look forward to the editor's response and the reviewers' comments.

Sincerely,

Mehmet Ergezer\\
Wentworth Institute of Technology

\end{document}
