\documentclass[10pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=0.72in]{geometry}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\usepackage{microtype}
\usepackage{enumitem}

\setlist[itemize]{leftmargin=1.4em,itemsep=0.12em,topsep=0.2em}
\setlength{\parskip}{0.32em}
\setlength{\parindent}{0pt}
\pagestyle{empty}

\begin{document}

\begin{flushright}
Mehmet Ergezer\\
Wentworth Institute of Technology\\
550 Huntington Avenue\\
Boston, MA 02115, USA\\
\today
\end{flushright}

Editors\\
\textit{Constraints}

Dear Editors,

Please consider the manuscript
\textit{``Salem--Spencer sets as a constraint-solving benchmark:
decomposition, certification, and split-policy effects at $r_3(212)$''}
for publication in \textit{Constraints}.

The manuscript studies a natural finite-domain feasibility family at the
frontier of OEIS A003002. It does not claim to resolve $r_3(212)$. Instead,
it develops and evaluates a reproducible constraint-solving methodology that
combines implied window-cardinality constraints, deterministic cube
decomposition, controlled solver comparisons, and independently checkable
infeasibility certificates.

The main contributions for the constraint-programming community are:

\begin{itemize}
\item A witness-informed CP-SAT architecture in which valid
  window-cardinality constraints reduce the controlled unresolved rate by
  about 28 percentage points, substantially more than the generic search
  interventions tested.
\item A matched five-policy split ablation and independent exhaustive cover
  audit that quantify a cover-size versus conquer-cost tradeoff. Global
  AP-degree reduces the complete depth-24 cover from 12,582,912 to 96,847
  cubes, but produces harder survivors under the historical CP-SAT cap.
\item Controlled configuration-level comparisons on released formulas.
  Native CaDiCaL 3.0.0 and kissat 4.0.4 close all 20 hard-pocket instances
  on matched hardware; a two-stage CDCL survey closes 6,045 of 6,071 broad
  residuals and leaves a compact 26-instance challenge tier.
\item A certificate and regression pipeline. Selected closures yield LRAT
  certificates accepted by the formally verified \texttt{cake\_lpr}
  checker, while end-to-end tests recover the known exact values
  $r_3(80)=22$, $r_3(90)=24$, and $r_3(100)=27$ from independently checked
  lower and upper evidence.
\end{itemize}

The released benchmark supports research on decomposition, problem-specific
implied constraints, search configuration, empirical hardness, and auditable
constraint solving. It includes deterministic generators, tiered JSONL and
CNF instances, matched solver outputs, model and cover audits, proof objects,
hashes, and machine-readable provenance.

Source code and compact artifacts are available in the
\href{https://github.com/memoatwit/erdos_1194/tree/constraints-submission-v1.5/erdos_r3}{tagged GitHub snapshot}.
The versioned dataset is archived on
\href{https://doi.org/10.5281/zenodo.21413746}{Zenodo}, and the preprint is
available as \href{https://arxiv.org/abs/2606.04016}{arXiv:2606.04016}.
\texttt{ARTIFACT\_REPRODUCIBILITY.md} maps the principal
claims to their data and reproduction commands; the $N=80$ exact-value
regression is the fastest complete check and runs in minutes on a workstation.
The manuscript is not under consideration elsewhere.

Generative-AI use is disclosed in the manuscript. The author made all final
scientific and experimental decisions, reviewed the generated text and code,
independently checked the reported claims, and takes full responsibility for
the work. The author declares no competing interests.

Thank you for your consideration.

Sincerely,

\vspace{0.5em}
Mehmet Ergezer\\
Wentworth Institute of Technology\\
\href{mailto:ergezerm@wit.edu}{ergezerm@wit.edu}\\
ORCID: \href{https://orcid.org/0000-0001-6627-3667}{0000-0001-6627-3667}

\end{document}
